---
title: "A7"
output:
  html_document:
    df_print: paged
    
header-includes:
  - \usepackage{bm}
  - \newcommand{\E}{\mathbb{E}}
  - \newcommand{\R}{\mathbb{R}}
  - \newcommand{\argmin}{\mathrm{arg\,min\;}}
  - \newcommand{\rank}{\mathrm{rank}}
  - \newcommand\independent{\protect\mathpalette{\protect\independenT}{\perp}}\def\independenT#1#2{\mathrel{\rlap{$#1#2$}\mkern2mu{#1#2}}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(dplyr)
```


```{r}
studentized_Tstar <- function(Xstar,X){
return( (mean(Xstar)-mean(X))/sd(Xstar)*sqrt(N))
}

non_studentized_Tstar <- function(Xstar,X){
return( (mean(Xstar)-mean(X))*sqrt(N))
}

sample_truth_scaled_Tstar <- function(Xstar,X){
return( (mean(Xstar)-mean(X))/sd(X)*sqrt(N))
}

```



```{r}
lambda <- 2
N <- 100
B <- 1000
asymptotic_frequency <- 0
studentized_frequency <- 0
non_studentized_frequency <- 0
sample_truth_scaled_frequency <- 0


for (i in 1:1000){
  X <- rexp(N,lambda)
  
  studentized_Tstars <- rep(0,B)
  non_studentized_Tstars <- rep(0,B)
  sample_truth_scaled_Tstars <- rep(0,B)
  
  
  #Generate a bootsrap
  boot_ind <- sample(1:N, size=N*B, replace=T)
  boot_data <- matrix(X[boot_ind],ncol=B)
  
  for (i in 1:B){
    studentized_Tstars[i] <- studentized_Tstar(boot_data[,i],X)
    non_studentized_Tstars[i] <- non_studentized_Tstar(boot_data[,i],X)
    sample_truth_scaled_Tstars[i] <- sample_truth_scaled_Tstar(boot_data[,i],X)
  }
  
  
  #calculate asymptotic confidence interval
  CI_asymptotic <- mean(X) + qnorm(0.95)*sd(X)/sqrt(N)
  if (CI_asymptotic > 0.5){
    asymptotic_frequency <- asymptotic_frequency + 1
  }
  
  #calculate studentized (bootstrap) confidence interval
  CI_studentized <- mean(X) + quantile(studentized_Tstars,0.95)*sd(X)/sqrt(N)
  if (CI_studentized > 0.5){
    studentized_frequency <- studentized_frequency + 1
  }
  
  #calculate non_studentized confidence interval
  CI_non_studentized <- mean(X) + quantile(non_studentized_Tstars,0.95)/sqrt(N)
  if (CI_non_studentized > 0.5){
    non_studentized_frequency <- non_studentized_frequency + 1
  }
  
  #calculate sample_truth_scaled confidence interval
  CI_sample_truth_scaled <- mean(X) + quantile(sample_truth_scaled_Tstars,0.95)*sd(X)/sqrt(N)
  if (CI_sample_truth_scaled > 0.5){
    sample_truth_scaled_frequency <- sample_truth_scaled_frequency + 1
  }
  
  
}

```


For this assignment, for $X_1,\ldots,X_{100}$ independently retrieved from $Exp(2)$, I considered the following CIs for $\E X_1 = 1/2$:

* asymptotic: $\left(-\infty, \bar{X}_N + \frac{\widehat{\sigma}}{\sqrt{N}} z(\alpha) \right)$
* studentized (bootstrap): $\left(-\infty, \bar{X}_N + \frac{\widehat{\sigma}}{\sqrt{N}} q^\star(\alpha) \right)$ with $T^\star = \sqrt{N}\frac{\bar{X}_N^\star - \bar{X}_N}{\widehat{\sigma}^\star}$
* non-studentized ($\star$): $\left(-\infty, \bar{X}_N + \frac{1}{\sqrt{N}} q^\star(\alpha) \right)$ with $T^\star = \sqrt{N}\big(\bar{X}_N^\star - \bar{X}_N\big)$
* sample-truth-scaled ($\star$): $\left(-\infty, \bar{X}_N + \frac{\widehat{\sigma}}{\sqrt{N}} q^\star(\alpha) \right)$ with $T^\star = \sqrt{N}\frac{\bar{X}_N^\star - \bar{X}_N}{\widehat{\sigma}}$

For each interval, I verified the coverage of the interval via a simulation study of 1000 runs. Below is the report of the coverage proportions of each interval.

From the table, we can see that the coverage of all the intervals are similar (90%-95%). The studentized interval had the lowest coverage, with around 90%, while the rest of the intervals have similar coverages of around 93%.


```{r}

freq_data <- tibble(Asymptotic=asymptotic_frequency/1000, Studentized=studentized_frequency/1000, Non_studentized=non_studentized_frequency/1000, Sample_truth_scaled=sample_truth_scaled_frequency/1000)

print(freq_data)

```
