---
title: "Small_Project_Markdown"
output: html_document
date: "2022-10-23"
author: "Issam Arabi, Emmanuele Sorgente, Eric Maeder"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

# Exploratory Data Analysis of Movies and Shows #

### Description ###
This project aims to analyse a data set concerning popular movies using exploratory tools.
The said data set initially contains information about 45'466 different shows, organized into the following co-variates. In brackets are the variables we do not treat :

+ The show original budget
+ The genre categories, i.e. Action, Adventure, ...
+ The original language the show was created in
+ The title
+ (a quick overview of the plot)
+ A numeric index indicating the show popularity
+ The production companies
+ (The production country)
+ The original release date of the show
+ The total revenue
+ The show duration in minutes
+ A variable named "vote_average" representing the mean note given by users on an integer scale from 1 to 10.
+ The total number of said votes

First, we decided to work the data set a bit to adjust the variable types, add interesting variables that are combinations of preexisting ones, and to remove subjectively uninteresting observations. Precisely, we removed from our data movies that have a small (<2000) or unknown budget, are under 10min long, or have under ten votes. Left are 1580 movies.

```{r rework, echo = FALSE}
library(ggplot2)


movies_metadata <- read.csv('./movies_metadata.csv') ### update path depending on how you stored it.
movies_metadata <- movies_metadata[,!names(movies_metadata) %in% c("adult", "belongs_to_collection", "homepage", 'id', 'imdb_id', 'poster_path', 'spoken_languages', 'status', 'tagline', "title", 'video')]
h=c()
for (i in (1:length(movies_metadata$budget))){
  if (isTRUE(movies_metadata$budget[i]<60000)==TRUE || is.na(movies_metadata$budget[i])==TRUE ||  isTRUE(movies_metadata$runtime[i]<10)==TRUE ||  isTRUE(movies_metadata$vote_count[i]<10)==TRUE){
    h=c(h,i)
  }### why remove low revenue or low vote average ?
}
movies_metadata=movies_metadata[-h,] ### retains only interesting variables.h contains some specific subset based on revenue ?
#length(movies_metadata$budget)

### Adjustment of the variable types:
movies_metadata$budget <- as.numeric(movies_metadata$budget)
movies_metadata$popularity <- as.numeric(movies_metadata$popularity)

```

New variables we deemed interesting to add:

+ a binary indicator of whether the show had a low or high production cost (threshold 10'000'000)
+ another binary covariate returning 'true' if the show was originally produced in english.
+ a profit variable, calculated deducing the production cost from the revenue.
+ a ternary 'profitable' variable, with profit thresholds 0 and 90'000'000.
+ a numeric percentprofit covariate corresponding to the profit as a percentage of the  initial investment.
+ binary variables 'Action','Adventure',... indicating whether the show is in said film genres or not.

```{r adding, echo=FALSE}
### calculating and adding variables of interest
movies_metadata$cost=1:length(movies_metadata$budget)  ### creates a binary cost variable based on budget, threshold 10^7
for (i in 1:length(movies_metadata$budget)) {
  if (isTRUE(movies_metadata$budget[i]<=10000000)==TRUE){
  movies_metadata$cost[i]= 'LOW'  
  }
  else {
  movies_metadata$cost[i]= 'HIGH'  
  }
}
movies_metadata$english=1:length(movies_metadata$budget) ### creates a binary 'english-based' english variable
for (i in 1:length(movies_metadata$budget)) {
  if (isTRUE(movies_metadata$original_language[i]=='en')==TRUE){
    movies_metadata$english[i]= 'YES'  
  }
  else {
    movies_metadata$english[i]= 'NO'  
  }
}

movies_metadata$profit=1:length(movies_metadata$budget) ### adds a profit variable, dependent on budget and revenue
for (i in 1:length(movies_metadata$budget)) {
  movies_metadata$profit[i]=movies_metadata$revenue[i]-movies_metadata$budget[i]
}

movies_metadata$profitable=1:length(movies_metadata$budget) ### creates a ternary profitable variable with thresholds 9*10^7 and 0
for (i in 1:length(movies_metadata$budget)) {
  if (isTRUE(movies_metadata$profit[i]>=90000000)==TRUE){
    movies_metadata$profitable[i]= 'HUGE'  
  }
  else if (isTRUE(movies_metadata$profit[i]>0)==TRUE & isTRUE(movies_metadata$profit[i]<90000000)==TRUE){
    movies_metadata$profitable[i]= 'YES'  
  }
  else {
    movies_metadata$profitable[i]= 'NO'  
  }
}
### percentprofit variable
movies_metadata$percentprofit <- movies_metadata$profit/movies_metadata$budget

### genre variables 
j <- which(grepl("Action",movies_metadata$genres)) #contains indices of films that are tagged 'Action'.

### possible relevant tags : Action, Adventure, Thriller, War, Comedy, Horror, Thriller, Crime, Drama, Animation, Fantasy, Science fiction

movies_metadata$Action=1:length(movies_metadata$genres) ### adds an Action variable for example

movies_metadata$Action <- grepl("Action",movies_metadata$genres)
movies_metadata$Adventure <- grepl("Adventure",movies_metadata$genres)
movies_metadata$Comedy <- grepl("Comedy",movies_metadata$genres)
movies_metadata$Drama <- grepl("Drama",movies_metadata$genres)
movies_metadata$Thriller <- grepl("Thriller",movies_metadata$genres)
```

Moving on to the exploration part, we first investigated the language variable. 

```{r english, echo = FALSE, out.width="50%"}
par(mfrow=c(1,5))

ggplot(data = movies_metadata) + geom_bar(mapping = aes(x = original_language)) + ggtitle("Movies by Language")

```

After noticing that most movies were in English, we decided to compare profits of English movies to movies in other languages.


```{r, echo = FALSE, out.width="50%"}
par(mfrow=c(1,5))

ggplot(data = movies_metadata) + geom_bar(mapping = aes(x = original_language, fill = profitable)) + ggtitle("Movies by Language, with profit")

ggplot(data = movies_metadata) + geom_bar(mapping = aes(x = profitable)) + facet_wrap(~ english) + ggtitle("Profit in English Movies vs. Others")

```

Next, we were interested in exploring the ratings of movies; In particular, we wanted to check wether the ratings of movies had any effect on the profit generated.


```{r, echo = FALSE, out.width="50%"}
par(mfrow=c(1,5))

ggplot(data = movies_metadata, mapping = aes(x = vote_average, y = profit, color = english, shape = english)) + 
  geom_point() + 
  stat_smooth(method=lm) + 
  facet_wrap(~ english) + 
  ggtitle("Effect of Ratings on Profit in English Movies vs. Others")

#ggplot(data = movies_metadata, mapping = aes(x = vote_count, y = profit, color = english, shape = english)) +
  #geom_point() +
  #stat_smooth(method=lm) +
  #facet_wrap(~ english)
```

### Areas of Interest ###

After exploring the data, the following questions about the data have raised our curiosity:
1. How is the revenue generated by a movie related to its budget?
2. Movies of which languages have the highest rating?
3. Over time, how has the runtime of movies changed (i.e. Are movies getting longer or shorter?)
4. Which companies produce the most popular movies?